Kelp morphometric properties and its link to the hydrodynamic modelling

### 2018-08-01 chapter_1.tex
removed some sections. I am keeping the background literature with regards to biotic drivers and overall kelp ecology for now. I think it may be important for the reader to understand the complexity of temperate reefs.

---
title: "Kelp morphometric properties and its link to the hydrodynamic modelling"
author: "Ross Coppin"
date: "26 February 2017"
output:
  word_document:
    toc: yes
  pdf_document:
    highlight: zenburn
    number_sections: yes
    toc: yes
---

## Summary

Marine ecosystems are maintained by a variety of complex interactions between abiotic and biotic variables such as temperature, wave exposure, pH, competition, and processes such as top-down and bottom-up control, predator-prey relationships and phenology [@Doney2012]. These abiotic and biotic variables, the interactions between them, and the various ecological processes, ultimately determine the community composition and ecological functioning of all ecosystems. Climate directly and indirectly affects the way in which abiotic and biotic variables interact, but is often compounded by other impacts such as habitat destruction, pollution, and over-fishing [@Blamey2015]. Temperature and wave exposure have been recognised as important variables with regards to climate-driven changes within the ocean. In order to persist and survive within variable and changing environments, organisms must either migrate, adapt, or die.

@@

## Background

Seaweeds, browns in particular, are capable of adapting morphological characteristics to persist in changing and variable ocean environments. Changes in morphology have been shown to increase photosynthetic ability, enhance nutrient uptake, and reduce drag with regards to changes in temperature and wave exposure. Wave exposure has been shown to be an important driver of seaweed morphology, as the main mechanism of seaweed mortality is through the mechanism of dislodgment. Changing morphology, such as decreasing frond area or strengthening through stipe circumference increase/reduces drag and increases the probability of survival. However, locally adapted seaweed may still be dislodged in pulse disturbance events such as storms, and may raft far distances and wash up on beaches. An example would be *Ecklonia maxima* which is a conspicuous brown seaweed along the coastline and beaches around South Africa. Therefore, the beach-cast may not always originate from adjacent kelp populations but rather from other regions which the individual is adapted for. Therefore, the kelp morphology may act as a proxy for investigating coastal currents and changes thereof. The advances in ocean hydrodynamic modelling has made great progress and has been applied in a variety of ways. Therefore, this study will use advances in hydrodynamic modelling in combination with kelp morphological characteristics to investigate coastal currents. Once the model has been established it may be applied in other ways, such as investigating the transport of micorplastics along the South African coastline. The harms of microplastics to the marine environment has gained much traction in recent years, but research in South Africa is lacking. The coupling of kelp morphology and the ability to simulate hydrodynamic processes can greatly improve our understanding of transport pathways and likely locations of accumulation. This in turn may inform management decisions with regards to eliminating and managing marine pollution in South Africa.

@@

Temperature and wave exposure have been shown to be important drivers of seaweed distribution, physiological functioning, ability to recover, population dynamics and morphology. Mechanical forces generated by the hydrodynamic environment, in the form of sudden strong ocean currents or storms, between 10- 20 m s-1 with accelerations of 400 m s-2 [@Friedland1995] are the biggest threat to kelp survival. Kelps are able to rapidly adapt their morphological characteristics to reduce drag and avoid dislodgment [@Blanchette1997]. For example a study by @Koehletal2008 showed that transplanted *Nereocystis luetkeana* plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days. Another study by @FowlerWalker2006 tested for differences in morphology of *Ecklonia radiata* between wave-sheltered and wave-exposed sites and was a combination of *in situ* sampling and transplant of juvenile plants. The results showed that morphology differed between wave-sheltered and wave-exposed sites (thin thallus at sheltered sites and a narrow, thick thallus with a thick stipe at exposed sites), and was consistent with previous studies. Juveniles transplanted into wave exposed sites under went rapid morphological adaption, whilst the opposite was true for wave-sheltered sites which showed slower morphological adaption.

@@

Advances in numerical modelling has gained much traction in recent years and has been applied in a variety of ways with regards to ecological studies. For example, a study by @wang2009 used the Delft3D-Flow model to assess the hydraulic suitability of a stream as a spawing ground for the Chinese Sturgeon (*Acipenser sinensis*) in the Yangtze River. The authors calculated the horizontal mean vorticity which was used to assess the hydraulic environment of spawning ground. The flow field state was determined through model simulation and field-measured data used to validate the model. The results added to existing scientific database for spawning ground hydraulic environmental protection. Different numerical models can often be integrated to model across ecosystem levels. For example a study by @leon2003 used integrated physical (Delft3D hydrological model) and bio-chemical (Agricultural Non-point Source model) processes models to investigate the possible impact on the Lake Malawi water quality due to management actions performed at the watershed level.

Since wave energy is an important driver in marine ecosystems, particularly kelp, the advances in hydrodynamic modelling offer a new opportunity for multifactoral and quantitative approach to research in marine ecosystems. The Delf3D and SWAN models have been used successfully in previous studies regarding brine plume discharge, impacts of storms, affects of climate change on the hydrological environment etc. The models have not been designed for shallow environments (<6m) and therefore may not be suitable to model coastal hydrological environments. However these models may be adjusted to suit coastal waters if they are combined with a new numerical model which can be calibrated to suit these needs.

@@

#Kelp morphology drivers

#Light

Light is an important factor for kelp survival, however if light is limited or excessive this may negatively impact kelp survival or growth. Much of the past research into the role light plays into the functioning of kelp [@Bruhn1996);@Belseth2012]. For instance, solar ultraviolate radiation has been shown to aﬀect sub-canopy Ecklonia radiata sporophytes when the canopy of mature *Ecklonia radiata* was removed [@Wood 1987]. The sub-canopy sporophytes experienced tissue damage,photopigment destruction,reduced growthand decreased survivorship, thus inhibiting their settlement and survival [@Wood 1987]. Laboratory experiments revealed that the UV component of radiation, rather than intense radiation itself, was responsible for the eﬀects mentioned above. High light stress has negative eﬀects, such as photoinhibition and photo-damage on *Ecklonia cava* sporophytes [@Altamirano2004]. @Altamirano2004 found that *Ecklonia cava* is more vulnerable to light stress conditions, and less likely to recover under unfavourable conditions [@Altamirano2004]. @Bolton1985 showed that under sub-saturating irradiances and supra- optimal temperatures *Ecklonia maxima* to showed a decrease in reproductive rates and an increase in cell production. A interesting ﬁnding of this study was that despite the decrease in reproductive rates, the ﬁnal egge production per female was greater under these conditons. The authors interpreted this an ecological adaption that may increase survival rates under times of stress or non - ideal conditions [@Bolton1985].

#Nutrients

The importance of nutrients in the functioning of kelps is well understood [@Dayton1985; @Gaylord2012]. Dissolved nitrogen, and in particular nitrate, are important; however research has also placed emphasis on phosphate and other trace compounds for functioning of kelps [@Dayton1985]. Additionally, some kelps have the ability to store inorganic nitrogen in order to compensate for periods of low nutrient availability, which has been observed for Laminaria and Macrocystis [@Dayton1985; @Gaylord2012]. Nutrient stratiﬁcation is also an important factor, particularly for canopy type kelps. The concentration of nutrients at the surface is important to the functioning and maintenance of the canopy. For instance kelp canopies in California often deteriorate in the summer months when surface nitrate levels are low [@Jackson1977]. Water motion is important in the assimilation of nutrients from the water column, and kelps have been shown to adapt blade morphology in order to create more turbulence around the boundary layer of the frond to enhance nutrient assimilation (Wheeler 1980). Temperature has also been closely linked with nutrient concentrations. Nutrients are often in higher concentrations in the water column during low temperature events. This is often an indication of an “up-welling” event, which brings cold and nutrient rich waters from the bottom to the surface of the water column. However temperature can play a direct role in the uptake of nutrients through eﬀects on algal metabolism; however this may vary from species to species [@Raven1988]. However most studies investigate multiple characteristics in one study (i.e. temperature and light and water motion), while studies investigating the direct role of temperature in nutrient assimilation is lacking in marine algae is lacking.

#Depth

Depth does not aﬀect kelp ecosystems directly, however a change in depth often causes ﬂuctuations or changes in other environmental variables such as water motion, light and temperature. Water motion also decreases with depth, and some kelps better suited to deeper environments (*L. pallida*) replace those in the shallows (*E. maxima*) (Dayton 1985; Gerard 1982). The increase in depth can lead to a decrease in sunlight penetration, with some species better adapted for low-light conditions than others, such as (*L. pallida*). Temperature may also change along a depth gradient due to a reduction in sunlight penetration. (Dayton 1985; Gerard 1982). Therefore depth does not directly play a role in kelp functioning but may alter more inﬂuential factors such as light and water motion.

# Temperature

Temperature Biogeographical shifts of kelp species can already be seen in areas where ocean warming is occurring. In Western Australia an increase in temperature has caused a biogeographical shift in kelp and urchins, which in turn has caused many kelp forests to shift to the less productive algal turfs [@Kelly2005];a review of the literature shows a consensus that biogeographical distributions of marine species are largely temperature related [@Dayton1985; @Jeﬀree1994; @Schils2006; @Sorte2010; @Bolton2012]. Other examples include Norway and eastern North America where ocean warming has facilitated kelp die-oﬀ and an increase in herbivore grazing rates [@Dayton1985]. The increase in grazing rates and the colonization of other algal species into these areas make it almost impossible for kelps to recover [@Kelly2005]. The majority of kelp species are artic and temperate organisms, and the warming of ocean temperatures is expected to cause a poleward biogeographical shift of species [@Bolton2012]. There is evidence to suggest that South African kelp forests are expanding due to ocean cooling [@Bolton2012], possibly driven by an intensiﬁcation and increase in coastal upwelling [@Blamey2012, @Blamey2015]. In South Africa there has been a biogeographical shift eastward along the coast due to a change in inshore temperature regime, making South Africa no exception to changing ocean temperatures [@Bolton2012]. Macroalgae, such as kelps, can react to an increase in surface temperatures in one of three ways: they can migrate, adapt and die [@Biskup2014]. A study by @Biskup2014 investigated the functional response of two kelp species (*Laminaria ochroleuca* and *Saccorhiza polyschides*) to rising sea temperatures. The functional responses of Saccorhiza polyschides was measured for both the subtidal and intertidal habitats, to see what aﬀect non- optimal conditions (intertidal zone) had on the kelps [@Rinde2005]. The study found that Laminaria ochroleuca exhibited a poor ability to acclimatise and was dependent on the kelp’s life history traits [@Biskup2014]. Therefore annual kelp species are more likely to survive under non-ideal condition, and the intertidal Saccorhiza polyschides, compared to the subtidal, showed a higher physiological ﬂexibility to changing conditions [@Biskup2014]. This may be because the intertidal zone undergoes far more change than the subtidal and therefore kelps in the intertidal are forced to adapt to harsher conditions where ﬂuctuations in temperature, sunlight, turbidity and water motion are common. The eﬀects on temperature have also been investigated by @Wernberg2010. The study looked at resilience of kelp beds along a latitudinal temperature gradient. Kelp abundance is likely to decline with the predicted warming of ocean waters @Wernberg2010 and although kelps have the ability to acclimatize and adjust their metabolic performance, which in turn allows them to change their physiological performance to mitigate the seasonal ﬂuctuations in temperature, this acclimatization is done at a cost @Wernberg2010. It is therefore important to understand the ecological implications of this cost in order to better understand the future impacts of rising sea temperatures @Wernberg2010. Results from the study showed that kelp beds in the northern regions had a lower resilience to perturbations, such as physical and biological disturbance, disease, competition from invasive species and decreasing water quality. This was as a result of low abundance of recruits, reduced physiological responsiveness and reliance on mature adult kelp for net recruitment, growth of recruits and recovery of canopy loss @Wernberg2010. Even though @Wernberg2010 acknowledged that it was diﬃcult to exclude the inﬂuences of other environmental parameters (sunlight, nutrients and wave exposure), temperature was most likely the main environmental driver of these eﬀects. While kelps may be able to maintain populations in the warmer upper limits of their distributions, the cost of acclimation will make them more susceptible to other stressors such as disturbance from anthropogenic sources, disease, competition and decreasing water quality. Changes in temperature can also impact the distribution of kelps.

# Wave exposure

## Aims of research

@@

A) Determine if the hydrodynamic environment is the main driver of kelp morphology and if this is specific to a location
B) Simulate kelp rafting by means of a hydrodynamic modelling and calibrate this model with *in situ* beach-cast morphometric data.
C) Use the calibrated hydrodynamic model to investigate dispersal of microplastics along the west coast and south-west coast of South Africa.

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Kelp morphometric properties and its link to the hydrodynamic modelling},
            pdfauthor={Ross Coppin},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Kelp morphometric properties and its link to the hydrodynamic modelling}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Ross Coppin}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{26 February 2017}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\subsection{Summary}\label{summary}

Marine ecosystems are maintained by a variety of complex interactions
between abiotic and biotic variables such as temperature, wave exposure,
pH, competition, and processes such as top-down and bottom-up control,
predator-prey relationships and phenology {[}@Doney2012{]}. These
abiotic and biotic variables, the interactions between them, and the
various ecological processes, ultimately determine the community
composition and ecological functioning of all ecosystems. Climate
directly and indirectly affects the way in which abiotic and biotic
variables interact, but is often compounded by other impacts such as
habitat destruction, pollution, and over-fishing {[}@Blamey2015{]}.
Temperature and wave exposure have been recognised as important
variables with regards to climate-driven changes within the ocean. In
order to persist and survive within variable and changing environments,
organisms must either migrate, adapt, or die.

@@

\subsection{Background}\label{background}

Seaweeds, browns in particular, are capable of adapting morphological
characteristics to persist in changing and variable ocean environments.
Changes in morphology have been shown to increase photosynthetic
ability, enhance nutrient uptake, and reduce drag with regards to
changes in temperature and wave exposure. Wave exposure has been shown
to be an important driver of seaweed morphology, as the main mechanism
of seaweed mortality is through the mechanism of dislodgment. Changing
morphology, such as decreasing frond area or strengthening through stipe
circumference increase/reduces drag and increases the probability of
survival. However, locally adapted seaweed may still be dislodged in
pulse disturbance events such as storms, and may raft far distances and
wash up on beaches. An example would be \emph{Ecklonia maxima} which is
a conspicuous brown seaweed along the coastline and beaches around South
Africa. Therefore, the beach-cast may not always originate from adjacent
kelp populations but rather from other regions which the individual is
adapted for. Therefore, the kelp morphology may act as a proxy for
investigating coastal currents and changes thereof. The advances in
ocean hydrodynamic modelling has made great progress and has been
applied in a variety of ways. Therefore, this study will use advances in
hydrodynamic modelling in combination with kelp morphological
characteristics to investigate coastal currents. Once the model has been
established it may be applied in other ways, such as investigating the
transport of micorplastics along the South African coastline. The harms
of microplastics to the marine environment has gained much traction in
recent years, but research in South Africa is lacking. The coupling of
kelp morphology and the ability to simulate hydrodynamic processes can
greatly improve our understanding of transport pathways and likely
locations of accumulation. This in turn may inform management decisions
with regards to eliminating and managing marine pollution in South
Africa.

@@

Temperature and wave exposure have been shown to be important drivers of
seaweed distribution, physiological functioning, ability to recover,
population dynamics and morphology. Mechanical forces generated by the
hydrodynamic environment, in the form of sudden strong ocean currents or
storms, between 10- 20 m s-1 with accelerations of 400 m s-2
{[}@Friedland1995{]} are the biggest threat to kelp survival. Kelps are
able to rapidly adapt their morphological characteristics to reduce drag
and avoid dislodgment {[}@Blanchette1997{]}. For example a study by
@Koehletal2008 showed that transplanted \emph{Nereocystis luetkeana}
plants from a wave sheltered site to a wave exposed site changed their
morphology to flat blades and narrow laterals that are less prone to
drag forces in 4-5 days. Another study by @FowlerWalker2006 tested for
differences in morphology of \emph{Ecklonia radiata} between
wave-sheltered and wave-exposed sites and was a combination of \emph{in
situ} sampling and transplant of juvenile plants. The results showed
that morphology differed between wave-sheltered and wave-exposed sites
(thin thallus at sheltered sites and a narrow, thick thallus with a
thick stipe at exposed sites), and was consistent with previous studies.
Juveniles transplanted into wave exposed sites under went rapid
morphological adaption, whilst the opposite was true for wave-sheltered
sites which showed slower morphological adaption.

@@

Advances in numerical modelling has gained much traction in recent years
and has been applied in a variety of ways with regards to ecological
studies. For example, a study by @wang2009 used the Delft3D-Flow model
to assess the hydraulic suitability of a stream as a spawing ground for
the Chinese Sturgeon (\emph{Acipenser sinensis}) in the Yangtze River.
The authors calculated the horizontal mean vorticity which was used to
assess the hydraulic environment of spawning ground. The flow field
state was determined through model simulation and field-measured data
used to validate the model. The results added to existing scientific
database for spawning ground hydraulic environmental protection.
Different numerical models can often be integrated to model across
ecosystem levels. For example a study by @leon2003 used integrated
physical (Delft3D hydrological model) and bio-chemical (Agricultural
Non-point Source model) processes models to investigate the possible
impact on the Lake Malawi water quality due to management actions
performed at the watershed level.

Since wave energy is an important driver in marine ecosystems,
particularly kelp, the advances in hydrodynamic modelling offer a new
opportunity for multifactoral and quantitative approach to research in
marine ecosystems. The Delf3D and SWAN models have been used
successfully in previous studies regarding brine plume discharge,
impacts of storms, affects of climate change on the hydrological
environment etc. The models have not been designed for shallow
environments (\textless{}6m) and therefore may not be suitable to model
coastal hydrological environments. However these models may be adjusted
to suit coastal waters if they are combined with a new numerical model
which can be calibrated to suit these needs.

@@

\section{Kelp morphology drivers}\label{kelp-morphology-drivers}

\section{Light}\label{light}

Light is an important factor for kelp survival, however if light is
limited or excessive this may negatively impact kelp survival or growth.
Much of the past research into the role light plays into the functioning
of kelp {[}@Bruhn1996);@Belseth2012{]}. For instance, solar ultraviolate
radiation has been shown to aﬀect sub-canopy Ecklonia radiata
sporophytes when the canopy of mature \emph{Ecklonia radiata} was
removed {[}@Wood 1987{]}. The sub-canopy sporophytes experienced tissue
damage,photopigment destruction,reducedgrowthand decreased survivorship,
thus inhibiting their settlement and survival {[}@Wood 1987{]}.
Laboratory experiments revealed that the UV component of radiation,
rather than intense radiation itself, was responsible for the eﬀects
mentioned above. High light stress has negative eﬀects, such as
photoinhibition and photo-damage on \emph{Ecklonia cava} sporophytes
{[}@Altamirano2004{]}. @Altamirano2004 found that \emph{Ecklonia cava}
is more vulnerable to light stress conditions, and less likely to
recover under unfavourable conditions {[}@Altamirano2004{]}. @Bolton1985
showed that under sub-saturating irradiances and supra- optimal
temperatures \emph{Ecklonia maxima} to showed a decrease in reproductive
rates and an increase in cell production. A interesting ﬁnding of this
study was that despite the decrease in reproductive rates, the ﬁnal egge
production per female was greater under these conditons. The authors
interpreted this an ecological adaption that may increase survival rates
under times of stress or non - ideal conditions {[}@Bolton1985{]}. Light
is also important for survival and functioning of benthic algal species;
a study by @Toohey2004 investiagted the possible link of light and
thallus scour of the kelp \emph{Ecklonia radiata canopy} on the
associated foliose algal assemblage. The authors concluded that a
reduction in light, rather than thallus scour has an eﬀect on foliose
algal assemblage. Reduced light was associated with reduced diveristy
and an marked increase abundance in certian foliose species, which were
able to acclimiate more light under low irradiances compared to other
species {[}@Toohey2004{]}. Kelp may also compete intra-specﬁcally or
inter-speciﬁcally for light. For instance, @Clark2004 investiagted eﬀect
of canopy shading of \emph{Macrocystis pyrifera} and understorey canopy
shading of \emph{Pterygophora californica} individually and in
combination of benthic macro-algae assemblage. Under all treatments
there was a sigiﬁance change in macro-algal assemblage, with certian
species able to recover in abundance more readily than others. The
authors devided these species into two categories; ``shade-tolerant''
and ``light adapted'', and conlcuded that certian foliose species are
better adapted for diﬀerent light conidtions in accordance with canopy
cover. In South Africa, the eﬀect of \emph{Ecklonia maxima} harvesting
on sub-story kelp, kelp population structure and kelp recruitement was
investigated by @Rothman2006. The results showed no signiﬁcant diﬀerence
in the three factors mentioned earlier and @Rothman2006 concluded that
factors other than light play a role in structuring kelp communities.

\section{Nutrients}\label{nutrients}

Nutrients
Theimportanceofnutrientsinthefunctioningofkelpsiswellunderstood
{[}@Dayton1985; @Gaylord2012{]}. Dissolved nitrogen, and in particular
nitrate, are important; however research has also placed emphasis on
phosphate and other trace compounds for functioning of kelps
{[}@Dayton1985{]}. Additionally, some kelps have the ability to store
inorganic nitrogen in order to compensate for periods of low nutrient
availability, which has been observed for Laminaria and Macrocystis
{[}@Dayton1985; @Gaylord2012{]}. Nutrient stratiﬁcation is also an
important factor, particularly for canopy type kelps. The concentration
of nutrients at the surface is important to the functioning and
maintenance of the canopy. For instance kelp canopies in California
often deteriorate in the summer months when surface nitrate levels are
low {[}@Jackson1977{]}. Water motion is important in the assimilation of
nutrients from the water column, and kelps have been shown to adapt
blade morphology in order to create more turbulence around the boundary
layer of the frond to enhance nutrient assimilation (Wheeler 1980).
Temperature has also been closely linked with nutrient concentrations.
Nutrients are often in higher concentrations in the water column during
low temperature events. This is often an indication of an ``up-welling''
event, which brings cold and nutrient rich waters from the bottom to the
surface of the water column. However temperature can play a direct role
in the uptake of nutrients through eﬀects on algal metabolism; however
this may vary from species to species {[}@Raven1988{]}. However most
studies investigate multiple characteristics in one study
(i.e.~temperature and light and water motion), while studies
investigating the direct role of temperature in nutrient assimilation is
lacking in marine algae is lacking.

\section{Depth}\label{depth}

Depth does not aﬀect kelp ecosystems directly, however a change in depth
often causes ﬂuctuations or changes in other environmental variables
such as water motion, light and temperature. Water motion also decreases
with depth, and some kelps better suited to deeper environments
(\emph{L. pallida}) replace those in the shallows (\emph{E. maxima})
(Dayton 1985; Gerard 1982). The increase in depth can lead to a decrease
in sunlight penetration, with some species better adapted for low-light
conditions than others, such as (\emph{L. pallida}). Temperature may
also change along a depth gradient due to a reduction in sunlight
penetration. (Dayton 1985; Gerard 1982). Therefore depth does not
directly play a role in kelp functioning but may alter more inﬂuential
factors such as light and water motion.

\section{Sedimentation}\label{sedimentation}

Sedimentation and scour can be highly destructive to kelp forests with
most detrimental impacts on the development stages such as spores,
gametophytes and young plants (Dayton 1985), and has been linked to
seasonal trends relating to wind direction {[}@Moore1972{]}.
@Devinny1978 found that sedimentation greatly reduced spore settlement
and which in turn greatly reduces the probability of survival for
Macrocystis Pyrifera gametophytes. Some species seem more resistant to
the eﬀects of sedimentation such as Laminaria farlowii and Eisena aborea
compared to others such as Macrocystis Pyrifera (Dayton 1985). The
impacts of sedimentation is not limited to the sub-tidal but has also
been shown to eﬀect inter-tidal kelp as well. For example a study by
{[}@Schiel2006{]} investigated the eﬀects of sedimentation on the
post-settlement stages of habitat-forming seaweeds. The results showed
high mortality of germlings at sites that had high sedimentation, and
was further supported through laboratory tests. Grazing was found to be
negligible and mortality of gametophytes was attributed to
sedimentation, although its eﬀects did vary with molluscan grazing
{[}@Schiel2006{]}. The presence of kelps themselves can also eﬀect the
degree of sedimentation. For instance, @Wernberg2005 investigated the
ability of Ecklonia radiata (canopy forming species) to modify the
physical environment and the related eﬀects on associated foliose algae.
An increase in kelp densities was marked by a signiﬁcant reduction of
light and sedimentation, and found that sedimentation did not have a
signiﬁcant eﬀect on associated foliose algae {[}@Wernberg2005{]}.
Therefore it is often diﬃcult to separate the direct eﬀects of each of
the environmental variables as they are often related to one another.
However, sea surface temperatures is one of the main predictors of
marine biological systems, as it has been shown to relate to
physiological and ecological processes at all levels {[}@Andersen2013;
@Bolton1985; @Bolton2012; @Dayton1985{]}. Kelp forests are no exception,
with temperature being the main limiting factor {[}@Dayton 1985{]}.

\section{Temperature}\label{temperature}

Temperature Biogeographical shifts of kelp species can already be seen
in areas where ocean warming is occurring. In Western Australia an
increase in temperature has caused a biogeographical shift in kelp and
urchins, which in turn has caused many kelp forests to shift to the less
productive algal turfs {[}@Kelly2005{]};a review of the literature shows
a consensus that biogeographical distributions of marine species are
largely temperature related {[}@Dayton1985; @Jeﬀree1994; @Schils2006;
@Sorte2010; @Bolton2012{]}. Other examples include Norway and eastern
North America where ocean warming has facilitated kelp die-oﬀ and an
increase in herbivore grazing rates {[}@Dayton1985{]}. The increase in
grazing rates and the colonization of other algal species into these
areas make it almost impossible for kelps to recover {[}@Kelly2005{]}.
The majority of kelp species are artic and temperate organisms, and the
warming of ocean temperatures is expected to cause a poleward
biogeographical shift of species {[}@Bolton2012{]}. There is evidence to
suggest that South African kelp forests are expanding due to ocean
cooling {[}@Bolton2012{]}, possibly driven by an intensiﬁcation and
increase in coastal upwelling {[}@Blamey2012, @Blamey2015{]}. In South
Africa there has been a biogeographical shift eastward along the coast
due to a change in inshore temperature regime, making South Africa no
exception to changing ocean temperatures {[}@Bolton2012{]}. Macroalgae,
such as kelps, can react to an increase in surface temperatures in one
of three ways: they can migrate, adapt and die {[}@Biskup2014{]}. A
study by @Biskup2014 investigated the functional response of two kelp
species (\emph{Laminaria ochroleuca} and \emph{Saccorhiza polyschides})
to rising sea temperatures. The functional responses of Saccorhiza
polyschides was measured for both the subtidal and intertidal habitats,
to see what aﬀect non- optimal conditions (intertidal zone) had on the
kelps {[}@Rinde2005{]}. The study found that Laminaria ochroleuca
exhibited a poor ability to acclimatise and was dependent on the kelp's
life history traits {[}@Biskup2014{]}. Therefore annual kelp species are
more likely to survive under non-ideal condition, and the intertidal
Saccorhiza polyschides, compared to the subtidal, showed a higher
physiological ﬂexibility to changing conditions {[}@Biskup2014{]}. This
may be because the intertidal zone undergoes far more change than the
subtidal and therefore kelps in the intertidal are forced to adapt to
harsher conditions where ﬂuctuations in temperature, sunlight, turbidity
and water motion are common. The eﬀects on temperature have also been
investigated by @Wernberg2010. The study looked at resilience of kelp
beds along a latitudinal temperature gradient. Kelp abundance is likely
to decline with the predicted warming of ocean waters @Wernberg2010 and
although kelps have the ability to acclimatize and adjust their
metabolic performance, which in turn allows them to change their
physiological performance to mitigate the seasonal ﬂuctuations in
temperature, this acclimatization is done at a cost @Wernberg2010. It is
therefore important to understand the ecological implications of this
cost in order to better understand the future impacts of rising sea
temperatures @Wernberg2010. Results from the study showed that kelp beds
in the northern regions had a lower resilience to perturbations, such as
physical and biological disturbance, disease, competition from invasive
species and decreasing water quality. This was as a result of low
abundance of recruits, reduced physiological responsiveness and reliance
on mature adult kelp for net recruitment, growth of recruits and
recovery of canopy loss @Wernberg2010. Even though @Wernberg2010
acknowledged that it was diﬃcult to exclude the inﬂuences of other
environmental parameters (sunlight, nutrients and wave exposure),
temperature was most likely the main environmental driver of these
eﬀects. While kelps may be able to maintain populations in the warmer
upper limits of their distributions, the cost of acclimation will make
them more susceptible to other stressors such as disturbance from
anthropogenic sources, disease, competition and decreasing water
quality. Changes in temperature can also impact the distribution of
kelps.

\subsection{Aims of research}\label{aims-of-research}

@@

\begin{enumerate}
\def\labelenumi{\Alph{enumi})}
\tightlist
\item
  Determine if the hydrodynamic environment is the main driver of kelp
  morphology and if this is specific to a location
\item
  Simulate kelp rafting by means of a hydrodynamic modelling and
  calibrate this model with \emph{in situ} beach-cast morphometric data.
\item
  Use the calibrated hydrodynamic model to investigate dispersal of
  microplastics along the west coast and south-west coast of South
  Africa.
\end{enumerate}

\end{document}
